Appendix to 'cubble: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-17 jss/appendix.Rmd
Add scripts for examples

>>>>

# Scripts for creating the example data

This section contains the codes for extracting the data for the examples discussed in the main manuscript.

## Historical maximum temperature

The script below presents the codes required to obtain the data `historical_tmax` used in the example of Section 5.2 *Australian historical maximum temperature*. The function `rnoaa::meteo_pull_monitors()` may take a while to query a large number of stations in the first time. A copy of the data is provided in the data folder of the paper repository at: https://github.com/huizezhang-sherry/paper-cubble.

```{r eval = FALSE}
library(tidyverse)
library(cubble)
all_stations <- rnoaa::ghcnd_stations() %>%
  filter(str_starts(id, "ASN")) %>% # Australian stations start wiht "ASN"
  filter(last_year >= 2020) %>%
  mutate(wmo_id = as.numeric(wmo_id), name = str_to_lower(name)) %>%
  select(-state, -gsn_flag) %>%
  select(id, longitude, latitude, elevation, name,
         wmo_id, element, first_year, last_year) %>%
  rename(long = longitude, lat = latitude, elev = elevation)

tmax_stations <- all_stations %>% 
  filter(element == "TMAX", first_year < 1970, !is.na(wmo_id))

raw_tmax <- all_stations %>%
  rowwise() %>%
  mutate(ts = list(rnoaa::meteo_pull_monitors(
    monitors = id, var = "TMAX",
    date_min = glue::glue("{first_year}-01-01"),
    date_max = glue::glue("{last_year}-12-31")
    ) %>%
      select(-id)
    )
  )

historical_tmax <- raw_tmax %>%
  select(-element) %>%
  unnest(ts) %>%
  mutate(tmax = tmax/10) %>%
  filter(lubridate::year(date) %in% c(1971: 1975, 2016:2020)) %>% 
  as_cubble(index = date, key = id, coords = c(long, lat))

save(historical_tmax, file = here::here("data/historical_tmax.rda"))
```

## Australian 2016-2020 climate data

The data `climate_full` used in the example in Section 5.3, 5.4, and 5.6 of the main paper can be obtained in a similar fashion with a slight change on the variable and date parameter in `rnoaa::meteo_pull_monitors()`. The full script is provided below and a copy of the data is also available in the data folder of the paper repository linked above.

```{r eval = FALSE}
# all the Australian stations have all of the three PRCP, TMAX, and TMIN recorded
aus_stations <- all_stations %>%
  filter(element %in% c("PRCP", "TMAX", "TMIN")) %>%
  nest(element: last_year) %>%
  rowwise() %>%
  filter(nrow(data) == 3) %>%
  select(-data)

aus_climate_raw <- aus_stations %>%
  rowwise() %>%
  mutate(ts = list(
    rnoaa::meteo_pull_monitors(
      monitors = id, var = c("PRCP", "TMAX", "TMIN"),
      date_min = "2016-01-01", date_max = "2020-12-31"
      ) %>%
      select(-id)
    )
  )

climate_full <- aus_climate_raw %>%
  unnest(ts) %>%
  mutate(tmax = tmax/10, tmin = tmin/10) %>%
  cubble::as_cubble(key = id, index = date, coords = c(long, lat))

save(climate_full, file = here::here("data/climate_full.rda"))
```
